Last glacial cycle glacier erosion potential in the Alps

### 2021-01-25 articles/erosion/alps-erosion.tex
Rework the methods. Looking good.

    The ice-sheet model set-up was presented, and the results discussed, in an
    earlier publication \citep{Seguinot.etal.2018}. The simulations use the
    Parallel Ice Sheet Model (PISM, development version~e9d2d1f), an open
    source, finite difference, shallow ice sheet model
    \citep{PISM-authors.2017}. The model includes temperature and water-content
    dependent creep, pseudo-plastic basal sliding accounting for till
    dilatation under high water pressure, bedrock deformation under the ice
    load, and a~positive degree-day (PDD) surface mass balance model. The model
    is initialized with assumed present-day ice thickness and equilibrium
    ice and bedrock temperature at 120\,ka, and ran to the present.
    Climate forcing is provided by a~monthly climatology from interpolated
    observational data \citep[WorldClim;][]{Hijmans.etal.2005} and the European
    Centre for Medium-Range Weather Forecasts Reanalysis Interim
    \citep[ERA-Interim;][]{Dee.etal.2011}, amended with temperature lapse-rate
    corrections from the European Project for Ice Coring in Antarctica
    \citep[EPICA;][] {Jouzel.etal.2007}, and in some cases, time-dependent
    palaeo-precipitation reductions. Lower-resolution, alternative climate
    scenarios are analyzed in the discussion part
    (Sect.~\ref{sec:sensitivity}).

    Physical parametres are listed in the earlier open-access publication
    \citep{Seguinot.etal.2018} and complete PISM set-up is stored in long-term
    archived model output metadata for this \todo{DOI}, and for the earlier
    \todo{DOIs} publications.

@@

    Modelled erosion rates, $\dot{e}$, are calculated from the modelled basal
    velocities, $u_\mathrm{b}$, using a empirical erosion power-law,
    %
    \begin{equation}
        \dot{e} = K_\mathrm{g} u_\mathrm{b}^l ,
    \end{equation}
    %
    where $K_g = 5.2\times 10^{-11}\,m^{1-l}\,a^{l-1}$ and $l = 2.34$ are
    empirical constants calibrated on quantified sediment yields and velocities
    from 15~Patagonian and Antarctic Peninsula tidewater glaciers
    \citep{Koppes.etal.2015}. It is important to note that there are both large
    uncertainties surrounding these numbers on the aforementioned glaciers
    \todo{quantify}, and little insurance that this erosion law applies
    elsewhere. Alternative erosion laws are analyzed in the discussion part
    (Sect.~\ref{sec:powerlaws}).

    \note{I initially built the paper around the erosion law by
          \citep{Herman.etal.2015}, that yielded similar erosion patterns but
          up to several km of cumulative erosion in the valleys.}

    Unlike landscape evolution model studies \aref, this study does not account
    for the feedback of glacier erosion onto bedrock topography and ice dynamics.
    Neither are other forms of erosion accounted for. Instead, the
    time-integrated erosion rate,
    %
    \begin{equation}
        e =  K_\mathrm{g} \int u_\mathrm{b}^l dt,
    \end{equation}
    %
    is later referred to as the cumulative ``erosion potential''. This erosion
    potential is numerically approximated using a time-step of 10\,a in the
    main simulation, and 100\,a in the alternative climate scenarios runs. It
    can already be noticed from the above formula, that with an exponent, $l$,
    higher than one, the cumulative erosion potential has a greater dependency
    on the basal ice dynamics, $u_\mathrm{b}$, than the timing of glacier
    cover, $dt$.

>>>>

    The ice-sheet model set-up was presented in an earlier publication
    \citep{Seguinot.etal.2018} and is briefly summarized here. The simulations
    were conducted using the Parallel Ice Sheet Model (PISM, development
    version~e9d2d1f), an open-source, finite-difference, shallow-ice and
    shallow-shelf model
    \citep{PISM-authors.2017}. The model includes temperature and water-content
    dependent creep, pseudo-plastic basal sliding accounting for till
    dilatation under high water pressure, bedrock deformation under the ice
    load, and a~positive degree-day (PDD) surface mass balance model. The model
    is initialized with assumed present-day ice thickness and equilibrium
    ice and bedrock temperature at 120\,ka, and ran to the present.
    The ice-sheet model physical parameters are listed in the earlier
    open-access publication \citep{Seguinot.etal.2018}, and the complete PISM
    set-up is stored in long-term archived model output metadata for the
    present study \todo{upload}, and for the earlier publication
    \citep{Seguinot.2020, Seguinot.2020a}.

    Climate forcing is provided by a~monthly climatology from interpolated
    observational data \citep[WorldClim;][]{Hijmans.etal.2005} and the European
    Centre for Medium-Range Weather Forecasts Reanalysis Interim
    \citep[ERA-Interim;][]{Dee.etal.2011}, amended with temperature lapse-rate
    corrections from the European Project for Ice Coring in Antarctica
    \citep[EPICA;][] {Jouzel.etal.2007}, and time-dependent paleo-precipitation
    reductions \citep{Huybrechts.2002}. Lower-resolution runs based on
    alternative paleoclimate scenarios \citep{Seguinot.etal.2018} are analyzed
    in the discussion part (Sect.~\ref{sec:sensitivity}).

@@

    Modelled erosion rates, $\dot{e}$, are calculated from the modelled basal
    velocities, $u_\mathrm{b}$, using a empirical erosion power-law,
    %
    \begin{equation}
        \dot{e} = K_\mathrm{g} u_\mathrm{b}^l ,
    \end{equation}
    %
    where $K_g = 5.2\times 10^{-11}\,m^{1-l}\,a^{l-1}$ and $l = 2.34$ are
    empirical constants calibrated on quantified fjord sediment yields and
    velocities from 15~Patagonian and Antarctic Peninsula tidewater glaciers
    \citep{Koppes.etal.2015}. Alternative erosion laws are analyzed in the
    discussion part (Sect.~\ref{sec:powerlaws}).

    Our approach does not account for the feedback of glacier erosion onto
    bedrock topography and ice dynamics. Neither are other forms of erosion
    accounted for. Instead, we compute the time-integrated erosion rate,
    %
    \begin{equation}
        e =  K_\mathrm{g} \int u_\mathrm{b}^l dt,
    \end{equation}
    %
    and refer to it as the cumulative ``erosion potential''. The integration
    is numerically approximated using a time-step of 10\,a in the
    main simulation, and 100\,a in the alternative climate scenarios runs. It
    can be noted from the above formula, that with an exponent, $l$,
    higher than one, the cumulative erosion potential has a greater dependency
    on the basal ice dynamics, $u_\mathrm{b}$, than the duration of glacier
    cover.
